### 2020-07-22 msr.tex
writing checklist and Kozlowski review comments

\glspl{MSR} are one of six advanced reactor designs shortlisted by
the Generation IV Forum in 2001 for promising significant advances in safety,
sustainability, efficiency, and cost over existing designs. This
has attracted significant attention and resources towards
\gls{MSR} research, most noticeable by the number of start-up companies that
have emerged in recent years touting various \gls{MSR} designs. This chapter
provides a brief history of \glspl{MSR}, followed by the distinctive features
that earned the concept the label of being a Generation IV reactor. Lastly,
this chapter presents the reference specifications of the \gls{MSFR} concept
studied in this work and summarizes the state of the art in the international
literature regarding modeling and simulation of the \gls{MSFR}.

@@

In spite of this setback, research into \glspl{MSR} continued through the late
1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
concept, called the \gls{DMSR} \cite{gehin_liquid_2016} with denatured
$^{235}$U fuel (i.e. low-enriched uranium). The \gls{ORNL} researchers
developed this design in response to the fuel reprocessing restrictions
introduced by President Ford in 1976. The \gls{DMSR} would operate as a
once-through
converter system without fuel reprocessing. While the fuel consists of 19.75
\% high-assay low-enriched uranium, the initial core loading includes thorium
to boost its conversion ratio throughout its lifetime. It has a continuous
online feed consisting of \gls{HALEU} to maintain criticality, and denatured
$^{235}$U to keep uranium enrichment levels below nuclear non-proliferation
policy thresholds. The design also includes a gas sparging system for removing
gaseous fission products, while noble metals plate out onto the walls of
the coolant loop. The older \gls{MSBR} design had a significant drawback; the
extensive neutron damage in the graphite moderator necessitated frequent
replacement (every four years) throughout its operational lifetime. The
\gls{DMSR} avoids this issue by having a lower power density while maintaining
the overall power output of 2250 MW$_{\text{th}}$. As a result, researchers
projected that the graphite moderator would last for the entirety of the
\gls{DMSR}'s design lifetime.

A concurrent program at the UK Atomic Energy Authority led the
development of a 2500 MW$_{\text{e}}$ lead-cooled Molten Chloride Fast Reactor
concept \cite{smith_assessment_1974}. It is a dual fluid system,
with separate loops for the fuel salt and the blanket salt. The blanket is a
1 m-wide tank surrounding the core. The absence of moderators and the choice
of chloride over fluoride salt resulted in a relatively hard neutron spectrum
which favors $^{239}$Pu breeding over the thorium cycle. The UK researchers
performed some experiments to study molten salt chemistry but there were no
reactor prototypes. The UK program eventually shut down just like its
US counterpart partly due to the successful demonstration of the Prototype
Fast Reactor which had achieved criticality in 1974.

Following a lull lasting through the late 20th century, researchers at
\gls{CNRS} began research into \glspl{MSR} in 1997
\cite{heuer_simulation_2010}. Starting from the \gls{MSBR} design, they
performed parametric studies based on safety, breeding, and
other performance metrics \cite{mathieu_thorium_2006}. Their efforts
culminated in the \gls{MSFR} concept, a fast-spectrum breeder \gls{MSR}
designed to run on the thorium fuel cycle \cite{merle_optimized_2007}. In
2008, the Generation IV International Forum highlighted the \gls{MSFR} among
other \gls{MSR} designs for further development \cite{gif_generation_2008}.
The \gls{MSFR} has also benefited from collaborative research through two
European-funded projects, the \gls{EVOL} \cite{euratom_final_2015} and
\gls{SAMOFAR} \cite{kloosterman_20_2017} projects. Under the \gls{EVOL}
project, researchers further optimized the \gls{MSFR} design based on
neutronic and thermal-hydraulic safety analyses. They also diversified
start-up fuel options beyond the original $^{233}$U-based fuel composition to
include $^{235}$U, $^{239}$Pu, and \gls{TRU}, and established fuel
reprocessing schemes for the \gls{MSFR}. The \gls{SAMOFAR} project, which
stated approximately 2 years after the end of the \gls{EVOL} project,
supported more comprehensive safety assessments of the reactor and the
reprocessing plant, and funded a number of experiments for validation of the
\gls{MSFR}'s safety features. This thesis presents steady-state and
transient results of the \gls{MSFR} in Moltres. Therefore, Section
\ref{sec:msfr} describes the \gls{MSFR} design in further detail.

@@

The various \gls{MSR} designs under development today illustrate the
flexibility of this reactor concept. Graphite-moderated thermal-spectrum
\glspl{MSR} are typically straightforward \gls{LEU} burners or
$^{232}$Th/$^{233}$U iso-breeders/breeders, while epithermal- and
fast-spectrum \glspl{MSR} can operate as \gls{TRU} fuel burners or
$^{238}$U/$^{239}$Pu breeders. Breeder designs can be further categorized into
one- or two-fluid designs. Two-fluid designs feature separate blanket molten
salt mixtures that contain higher proportions of fertile material than the
fuel salt mixture. Examples of one-fluid designs include the Integral Molten
Salt Reactor \cite{leblanc_integral_2015} and the Transatomic Power
\gls{MSR} design \cite{transatomic_power_corporation_technical_2016} while
two-fluid designs include the \gls{MSFR} \cite{merle_optimized_2007} and
\gls{MOSART} \cite{ignatiev_molten_2014}.

@@

\glspl{MSR} rely on natural physical phenomena for passive safety such as the
strong negative temperature reactivity feedback of the fuel salt due to
greater temperature-induced expansion in liquid fuel than solid
fuel. Combined with the Doppler broadening of resonance capture cross sections
present in both fuel forms, there would be a smaller temperature
increase following an unprotected reactivity insertion. The overall
temperature reactivity coefficient varies widely among \gls{MSR}
designs due to other structures, such as moderators and reflectors, present in
the core. In particular, graphite moderators tend to have slightly positive
temperature reactivity coefficients. The \gls{MSBR} concept has this issue,
but the total temperature reactivity coefficient is still relatively large and
negative \cite{rykhlevskii_modeling_2019}. The negative temperature reactivity
feedback provides a great degree of control and stability as it is always
present in an \gls{MSR} regardless of the operating conditions.

@@

However, \glspl{MSR} do have some associated safety risks.
Firstly, \glspl{MSR} have smaller fractions of delayed neutron precursors in
the active core region as some of them decay in the external loop regions.
This complicates reactor control and may result in faster transients due to
the decrease in average neutron lifetime. Secondly, hot molten salts are
corrosive and the corrosion mechanism is different from conventional
corrosion induced by water and other common agents \cite{yoshioka_7_2017}. The
intense gamma and neutron radiation in the core may also accelerate corrosion
in the structural components. Lastly, overcooling in pipes and heat exchangers
may pose operational challenges as the salt can freeze and restrict flow,
causing a loss of flow accident \cite{ho_molten_2013}.

@@

In more recent years, there has been
significant progress towards fully coupled, spatial codes that feature
2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
\cite{cammi_multi-physics_2011} performed a ``\gls{MPM}'' analysis
of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
using the commercial finite element analysis software COMSOL Multiphysics. The
physics were implemented through the two-group neutron diffusion equations,
and the \gls{RANS} standard $k$-$\epsilon$ turbulence model, for the
neutronics and thermal-hydraulics respectively. The
authors emphasized the need for proper full coupling of the multiphysics and
presented both steady-state and transient results in various
scenarios such as reactivity insertions, changes in pumping rate, and the
presence of periodic perturbations. Nagy et al. \cite{nagy_steady-state_2014}
introduced an improved version of DALTON which coupled with a \gls{CFD}
application and incorporated incompressible
Navier-Stokes flow to simulate salt flow more accurately. The authors
validated their model against \gls{MSRE} experimental data before presenting
multiphysics simulation results of a different graphite-moderated breeder
\gls{MSR}.

Both DALTON and the COMSOL-based tool featured again in a 2014 paper by
Fiorina et al. \cite{fiorina_modelling_2014} for a 2-D axisymmetric
model of the \gls{MSFR}. The authors, from Politecnico di Milano and
Technical University of Delft, compared \gls{MSFR} transient safety
analysis results from both tools. They referred to the COMSOL and DALTON/HEAT
\gls{MSFR} models as the Polimi and TUDelft models, respectively, after the
universities where each application was developed. This thesis also follows
this naming convention. With
multigroup neutron diffusion and \gls{RANS} formulations on ultra fine meshes,
both models agreed on the steady-state neutron flux, temperature,
and \gls{DNP} distributions, as well as the power responses following various
accident transient initiations.
Aufiero et al. \cite{aufiero_development_2014} concurrently developed
a full-core 3-D model of the \gls{MSFR} in OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. Aufiero et al. also
provided quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.

@@

The \gls{MSFR} is a European reference fast-spectrum \gls{MSR} concept
\cite{mathieu_thorium_2006, merle_optimized_2007}. Table \ref{table:msfr} and
Figure \ref{fig:msfr} show the main specifications and schematic view of the
\gls{MSFR}, respectively. Developed from the \gls{MSBR} design, the
\gls{MSFR} is intended to run primarily on a closed thorium fuel cycle with
continuous online fuel reprocessing. Several reasons motivated the omission of
graphite moderators from the original \gls{MSBR} design. Firstly, graphite is
susceptible to
long-term radiation damage and replacement is likely to be necessary during
the operating lifetime of the reactor. Secondly, graphite also has a positive
temperature coefficient of reactivity; eliminating graphite from the design
ensures a greater safety margin \cite{mathieu_thorium_2006}. Lastly, while
negative temperature coefficients are attainable with very thermalized
spectra, breeding ratios deteriorated significantly due to parasitic
absorption in the graphite \cite{mathieu_thorium_2006}.
%
\begin{figure}[htb!] 
	\centering
	\includegraphics[width=0.8\textwidth]{MSFR}
	\caption{Schematic view of the MSFR concept. Figure reproduced from
	Brovchenko et al. \cite{brovchenko_design-related_2013}.}
	\label{fig:msfr}
\end{figure}
%
\begin{table}[htb!]
    \small
	\caption{Main specifications of the \gls{MSFR} concept
				\cite{merle_concept_2017}.}
	\centering
	\begin{tabular}{ l r }
		\toprule
		Parameter & Value \\
		\midrule
		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
		1300 
		\\
		Salt volume [m$^3$] & 18 \\
		Salt fraction in core & 0.5 \\
		Number of circulation loops & 16 \\
		Nominal flow rate [kg s$^{-1}$] & 18500  \\
		Nominal circulation time [s] & 4.0 \\
		Inlet/outlet temperature [K] & 923 / 1023 \\
		Blanket volume [m$^3$] & 7.3\\
		\bottomrule
	\end{tabular}
	\label{table:msfr}
\end{table}

In the \gls{MSFR} design, fuel salt flows vertically upward through a 9 m$^3$
central core region. At the top of the core, the flow separates into sixteen
smaller external loops, each of which passes through a heat exchanger before
being pumped back into the bottom of the core. The salt also passes through
the online salt reprocessing and gas sparging systems located along the
external loops. A toroidal blanket tank containing fertile salt for breeding
surrounds the core radially. The top and bottom of the core are enclosed
by nickel alloy reflectors. A layer of boron carbide behind the blanket tank
protects the peripheral equipment from excessive neutron damage. During severe
accidents when core temperatures rise to dangerous levels, the actively
fan-cooled freeze plug at the bottom of the core melts and drains the fuel
salt into a containment vessel designed to keep the salt subcritical.

Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
cycle, it supports a range of start-up fuel and feed compositions. This
versatility is particularly important for the first few \glspl{MSFR} to be
deployed due to the lack of $^{233}$U reserves required for the initial core
loading. In general, the fuel and blanket salts are approximately composed of
eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$; AcF$_4$ represents
actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
fluorides. Using fluoride salts instead of chloride salts results in a
relatively thermal neutron spectrum. Therefore, the \gls{MSFR} has a more
thermal neutron spectrum than a typical fast-spectrum reactors. For an initial
composition consisting of $^{232}$Th and $^{233}$U,
the benchmark value for the amount of $^{233}$U for criticality under
normal operating conditions is 2.5 mol\%. However, a neutronic benchmark
study by Brovchenko et al. \cite{brovchenko_neutronic_2019} shows that
different neutronics software with different nuclear data provide
different $k_{eff}$ estimates even with the same isotopic compositions and
temperature distributions. The individual co-authors adjusted the
ratio of $^{232}$Th to $^{233}$U slightly to achieve exact criticality at a
uniform temperature of 973 K for their own neutronics software
\cite{brovchenko_neutronic_2019}. This thesis
performs the same exercise to adjust the inlet and outlet temperatures to
match nominal values.

The thermal and electric power output of the \gls{MSFR} are 3000 MW
$_{\text{th}}$ and 1300 MW$_{\text{e}}$, respectively. The high thermal
efficiency ($\eta_{th}=0.43$) is due to the high operating
temperature. The inlet and outlet temperature
specifications of the fuel salt are 923 K and 1023 K, respectively, for a
minimum 50 K temperature buffer between the
operating temperatures and the melting point of the salt (873 K)
\cite{euratom_final_2015}. The \gls{MSFR} has heat exchangers and an
intermediate molten salt coolant loop to separate the power
conversion system from the highly radioactive fuel salt. This also serves as
a layer of containment between the radioactive material and the outside
environment. The exact composition of the intermediate coolant is not
finalized yet but potential candidates include NaF-NaBF$_4$, FLiNaK,
LiF-ZrF$_4$, and FliBe \cite{merle_concept_2017}.

@@

The present work uses the same 2-D square-cylindrical \gls{MSFR} design to
compare the present results to results from the Polimi and TUDelft models
\cite{fiorina_modelling_2014}. The design is a 2-D axisymmetric representation
of the \gls{MSFR} with the sixteen individual external loops homogenized into
a single outer loop as shown in Figure \ref{fig:msfrgeom}. For the multigroup
group constants calculations in Serpent, the 2-D axisymmetric model is
extended into a 3-D model by a 360-degree rotation about the
central axis. The material definitions are the same as those specified in the
reference \gls{MSFR} model \cite{brovchenko_neutronic_2019}. Accordingly, the
pump and heat exchanger regions
are assumed to be composed of 100\% fuel salt. While this is an approximation,
the outer loop is shielded from the core by a layer of boron carbide and sees
sufficiently low neutron fluxes that we may neglect its exact neutronic
properties.
%
\begin{figure}[t!] 
	\centering
	\includegraphics[width=0.5\textwidth]{msfr-geom}
	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
	simulations in Serpent. All dimensions are in meters.
	\cite{brovchenko_neutronic_2019}}
	\label{fig:msfrgeom}
\end{figure}

Although the present work uses the same 2-D axisymmetric geometry for
generating group constant data from Serpent, there are two minor differences
between the \gls{MSFR} geometry modeled herein, and the geometry used in the
Polimi and TUDelft models \cite{fiorina_modelling_2014}. First, the reactor
geometry for
Moltres excludes the 2-cm-thick structural material around the blanket tank
that separates the fuel and blanket salts. The thickness is much smaller than
the rest of the regions in the geometry and it complicated mesh generation in
the adjacent regions near the boundaries. The
neutronics results in Chapter \ref{chap:nts} show that the overall $k_{eff}$
and other parameters from Moltres show good agreement with that from Serpent.
It also has no direct impact on the temperature distribution results because
the present work solves for the temperature distribution in the primary loop
with homogeneous Neumann boundary conditions on the fuel salt-to-wall
interface. This approach is for consistency with the Polimi and TUDelft
models.

The second difference pertains to the out-of-core section of the primary loop.
The focus for Moltres development has been on core multiphysics over detailed
out-of-core multiphysics. The Moltres model simulates the outer loop as a 1-D
pipe with a pointwise heat sink to represent the heat exchanger. To simulate
the pumps, a Dirichlet boundary condition for velocity at the inlet boundaries
drives the flow in the central core region of the primary loop. At every
timestep, Moltres converges the core and outer loop calculations using Picard
iterations. This approach shares some similarities with the
geometric multiscale modeling approach by Zanetti et al.
\cite{zanetti_geometric_2015}. Future models could create a better
representation of the primary loop by implementing a whole continuous loop
with pressure increases and drops corresponding to the pumps and heat
exchangers.

>>>>

\glspl{MSR} are one of six advanced reactor classes shortlisted by
the Generation IV Forum in 2001 for promising significant advances in safety,
sustainability, efficiency, and cost over existing designs. This
attracted significant attention and resources towards
\gls{MSR} research, most noticeable by the number of start-up companies that
have emerged in recent years touting various \gls{MSR} designs. This chapter
provides a brief history of \glspl{MSR}, followed by the distinctive features
that earned the concept the label of being a Generation IV reactor. Lastly,
this chapter presents the reference specifications of the \gls{MSFR} concept
studied in this work and summarizes the state of the art in the international
literature regarding modeling and simulation of the \gls{MSFR}.

@@

In spite of this setback, research into \glspl{MSR} continued through the late
1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
concept, called the \gls{DMSR} \cite{gehin_liquid_2016} with denatured
$^{235}$U fuel (i.e. low-enriched uranium). The \gls{ORNL} researchers
developed this design in response to the fuel reprocessing restrictions
introduced by President Ford in 1976. The \gls{DMSR} would operate as a
once-through
converter system without fuel reprocessing. While the fuel consists of 19.75
\% high-assay low-enriched uranium, the initial core loading includes thorium
to boost its conversion ratio throughout its lifetime. It has a continuous
online feed consisting of \gls{HALEU} to maintain criticality, and denatured
$^{235}$U to keep uranium enrichment levels below nuclear non-proliferation
policy thresholds. The design also includes a gas sparging system for removing
gaseous fission products, while noble metals plate out onto the walls of
the coolant loop. The older \gls{MSBR} design had a significant drawback; the
extensive neutron damage in the graphite moderator necessitated frequent
replacement (every four years) throughout its operational lifetime. The
\gls{DMSR} avoids this issue by running on a lower power density while
maintaining
the overall power output of 2250 MW$_{\text{th}}$. As a result, researchers
projected that the graphite moderator would last for the entirety of the
\gls{DMSR}'s design lifetime.

A concurrent program at the UK Atomic Energy Authority led the
development of a 2500 MW$_{\text{e}}$ lead-cooled Molten Chloride Fast Reactor
concept \cite{smith_assessment_1974}. It is a dual fluid system,
with separate loops for the fuel salt and the blanket salt. The blanket is a
1 m-wide tank surrounding the core. The absence of moderators and the choice
of chloride over fluoride salt resulted in a relatively hard neutron spectrum
which favors $^{239}$Pu breeding over the thorium cycle. The UK researchers
performed some experiments to study molten salt chemistry but they did not
build reactor prototypes. The UK program eventually shut down just like its
US counterpart partly due to the successful demonstration of the Prototype
Fast Reactor which had achieved criticality in 1974.

Following a lull lasting through the late 20th century, researchers at
\gls{CNRS} began research into \glspl{MSR} in 1997
\cite{heuer_simulation_2010}. Starting from the \gls{MSBR} design, they
performed parametric studies based on safety, breeding, and
other performance metrics \cite{mathieu_thorium_2006}. Their efforts
culminated in the \gls{MSFR} concept, a fast-spectrum breeder \gls{MSR}
designed to run on the thorium fuel cycle \cite{merle_optimized_2007}. In
2008, the Generation IV International Forum highlighted the \gls{MSFR} among
other \gls{MSR} designs for further development \cite{gif_generation_2008}.
The \gls{MSFR} has also benefited from collaborative research through two
European-funded projects, the \gls{EVOL} \cite{euratom_final_2015} and
\gls{SAMOFAR} \cite{kloosterman_20_2017} projects. Under the \gls{EVOL}
project, researchers further optimized the \gls{MSFR} design based on
neutronic and thermal-hydraulic safety analyses. They also diversified
start-up fuel options beyond the original $^{233}$U-based fuel composition to
include $^{235}$U, $^{239}$Pu, and \gls{TRU}, and established fuel
reprocessing schemes for the \gls{MSFR}. The \gls{SAMOFAR} project, which
started approximately two years after the end of the \gls{EVOL} project,
supported more comprehensive safety assessments of the reactor and the
reprocessing plant, and funded a number of experiments for validation of the
\gls{MSFR}'s safety features. This thesis presents steady-state and
transient results of the \gls{MSFR} in Moltres. Therefore, Section
\ref{sec:msfr} describes the \gls{MSFR} design in further detail.

@@

The various \gls{MSR} designs under development today illustrate the
flexibility of this reactor concept. Graphite-moderated thermal-spectrum
\glspl{MSR} are typically straightforward \gls{LEU} burners or
$^{232}$Th/$^{233}$U breeders, while epithermal- and
fast-spectrum \glspl{MSR} can operate as \gls{TRU} fuel burners or
$^{238}$U/$^{239}$Pu breeders. Breeder designs can be further categorized into
one- or two-fluid designs. Two-fluid designs feature separate blanket molten
salt mixtures that contain higher proportions of fertile material than the
fuel salt mixture. Examples of one-fluid designs include the Integral Molten
Salt Reactor \cite{leblanc_integral_2015} and the Transatomic Power
\gls{MSR} design \cite{transatomic_power_corporation_technical_2016} while
two-fluid designs include the \gls{MSFR} \cite{merle_optimized_2007} and
\gls{MOSART} \cite{ignatiev_molten_2014}.

@@

\glspl{MSR} rely on natural physical phenomena for passive safety such as the
strong negative temperature reactivity feedback of the fuel salt due to
greater temperature-induced expansion in liquid fuel than solid
fuel. Combined with the Doppler broadening of resonance capture cross sections
present in both fuel forms, an \gls{MSR} would see a smaller temperature
increase following an unprotected reactivity insertion. The overall
temperature reactivity coefficient varies widely among \gls{MSR}
designs due to other structures, such as moderators and reflectors, present in
the core. In particular, graphite moderators tend to have slightly positive
temperature reactivity coefficients. The \gls{MSBR} concept has this issue,
but the total temperature reactivity coefficient is still relatively large and
negative \cite{rykhlevskii_modeling_2019}. The negative temperature reactivity
feedback provides a great degree of control and stability as it is always
present in an \gls{MSR} regardless of the operating conditions.

@@

However, \glspl{MSR} do have some associated safety risks.
Firstly, \glspl{MSR} have smaller fractions of delayed neutron precursors in
the active core region as some of the decay in the outer loop regions.
This complicates reactor control and may result in faster transients due to
the decrease in average neutron lifetime. Secondly, hot molten salts are
corrosive and the corrosion mechanism is different from conventional
corrosion induced by water and other common agents \cite{yoshioka_7_2017}. The
intense gamma and neutron radiation in the core may also accelerate corrosion
in the structural components. Lastly, overcooling in pipes and heat exchangers
may pose operational challenges as the salt can freeze and restrict flow,
causing a loss of flow accident \cite{ho_molten_2013}.

@@

In more recent years, researchers made significant progress towards fully
coupled, spatially resolved computational tools that feature
2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
\cite{cammi_multi-physics_2011} performed a ``Multi-Physics Modelling''
analysis
of a simplified 2-D axisymmetric model of a single \gls{MSBR} fuel channel
using the commercial finite element analysis software COMSOL Multiphysics. The
physics were implemented through the two-group neutron diffusion equations,
and the \gls{RANS} standard $k$-$\epsilon$ turbulence model, for the
neutronics and thermal-hydraulics respectively. The
authors emphasized the need for proper full coupling of the multiphysics and
presented both steady-state and transient results in various
scenarios such as reactivity insertions, changes in pumping rate, and the
presence of periodic perturbations. Nagy et al. \cite{nagy_steady-state_2014}
introduced an improved version of DALTON which coupled with a \gls{CFD}
application and incorporated incompressible
Navier-Stokes flow to simulate salt flow more accurately. The authors
validated their model against \gls{MSRE} experimental data and presented
multiphysics simulation results of a different graphite-moderated breeder
\gls{MSR}.

Fiorina et al. \cite{fiorina_modelling_2014} adopted both DALTON and the
COMSOL-based application in their study of a 2-D axisymmetric model of the
\gls{MSFR}. The authors, from Politecnico di Milano and Technical University
of Delft, compared \gls{MSFR} transient safety analysis results from both
tools. They referred to the COMSOL and DALTON/HEAT
\gls{MSFR} models as the Polimi and TUDelft models, respectively, after the
universities where each application was developed. This thesis also follows
this naming convention. With
multigroup neutron diffusion and \gls{RANS} formulations on ultra fine meshes,
both models agreed on the steady-state neutron flux, temperature,
and \gls{DNP} distributions, as well as the power responses following various
accident transient initiations.
Aufiero et al. \cite{aufiero_development_2014} concurrently developed
a full-core 3-D model of the \gls{MSFR} in OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. Aufiero et al. also
provided quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.

@@

The \gls{MSFR} is a European reference fast-spectrum \gls{MSR} concept
\cite{mathieu_thorium_2006, merle_optimized_2007}. Table \ref{table:msfr} and
Figure \ref{fig:msfr} show the main specifications and schematic view of the
\gls{MSFR}, respectively. Developed from the \gls{MSBR} design, the
\gls{MSFR} is intended to run primarily on a closed thorium fuel cycle with
continuous online fuel reprocessing. Several reasons motivated the omission of
graphite moderators from the original \gls{MSBR} design. Firstly, graphite is
susceptible to
long-term radiation damage and replacement is likely to be necessary during
the operating lifetime of the reactor. Secondly, graphite also has a positive
temperature coefficient of reactivity; eliminating graphite from the design
ensures a greater safety margin \cite{mathieu_thorium_2006}. Lastly, while
negative temperature coefficients are attainable with extremely thermalized
spectra, these configurations require large volumes of graphite which
significantly deteriorate breeding ratios due to parasitic absorption
\cite{mathieu_thorium_2006}.
%
\begin{figure}[htb!] 
	\centering
	\includegraphics[width=0.7\textwidth]{MSFR}
	\caption{Schematic view of the MSFR concept. Figure reproduced from
	Brovchenko et al. \cite{brovchenko_design-related_2013}.}
	\label{fig:msfr}
\end{figure}
%
\begin{table}[htb!]
    \small
	\caption{Main specifications of the \gls{MSFR} concept
				\cite{merle_concept_2017}.}
	\centering
	\begin{tabular}{ l r }
		\toprule
		Parameter & Value \\
		\midrule
		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
		1300 
		\\
		Salt volume [m$^3$] & 18 \\
		Salt fraction in core & 0.5 \\
		Number of circulation loops & 16 \\
		Nominal flow rate [kg s$^{-1}$] & 18500  \\
		Nominal circulation time [s] & 4.0 \\
		Inlet/outlet temperature [K] & 923 / 1023 \\
		Blanket volume [m$^3$] & 7.3\\
		\bottomrule
	\end{tabular}
	\label{table:msfr}
\end{table}

In the \gls{MSFR} design, fuel salt flows vertically upward through a 9 m$^3$
central core region. At the top of the core, the flow separates into sixteen
smaller outer loops, each of which passes through a heat exchanger before
being pumped back into the bottom of the core. The salt also passes through
the online salt reprocessing and gas sparging systems located along the
outer loops. A toroidal blanket tank containing fertile salt for breeding
surrounds the core radially. The top and bottom of the core are enclosed
by nickel alloy reflectors. A layer of boron carbide behind the blanket tank
protects the peripheral equipment from excessive neutron damage. During severe
accidents when core temperatures rise to dangerous levels, the actively
fan-cooled freeze plug at the bottom of the core melts and drains the fuel
salt into a containment vessel designed to keep the salt subcritical.

Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
cycle, it supports a range of start-up fuel and feed compositions. This
versatility is particularly important for the first few \glspl{MSFR} to be
deployed due to the lack of $^{233}$U reserves required for the initial core
loading. In general, the fuel and blanket salts are approximately composed of
eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$; AcF$_4$ represents
actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
fluorides. Using fluoride salts instead of chloride salts results in a
relatively thermal neutron spectrum. Therefore, the \gls{MSFR} has a more
thermal neutron spectrum than a typical fast-spectrum reactors. For an initial
composition consisting of $^{232}$Th and $^{233}$U,
the benchmark value for the amount of $^{233}$U for criticality under
normal operating conditions is 2.5 mol\%. However, a neutronic benchmark
study by Brovchenko et al. \cite{brovchenko_neutronic_2019} shows that
different neutronics software with different nuclear data provide
different $k_{eff}$ estimates even with the same isotopic compositions and
temperature distributions. The individual co-authors adjusted the
ratio of $^{232}$Th to $^{233}$U slightly to achieve exact criticality at a
uniform temperature of 973 K for their neutronics software
\cite{brovchenko_neutronic_2019}. This thesis
performs the same exercise to adjust the inlet and outlet temperatures to
match nominal values.

The thermal and electric power outputs of the \gls{MSFR} are 3000 MW
$_{\text{th}}$ and 1300 MW$_{\text{e}}$, respectively. The high thermal
efficiency ($\eta_{th}=0.43$) is due to the high operating
temperature. The inlet and outlet temperature
specifications of the fuel salt are 923 K and 1023 K, respectively, for a
minimum 50 K temperature buffer between the
operating temperatures and the melting point of the salt (873 K)
\cite{euratom_final_2015}. The \gls{MSFR} has heat exchangers and an
intermediate molten salt coolant loop to separate the power
conversion system from the highly radioactive fuel salt. This also serves as
a layer of containment between the radioactive material and the outside
environment. The exact composition of the intermediate coolant is not
finalized yet but potential candidates include NaF-NaBF$_4$, FLiNaK,
LiF-ZrF$_4$, and FliBe \cite{merle_concept_2017}.

@@

The present work uses the same 2-D square-cylindrical \gls{MSFR} design to
compare the present results to results from the Polimi and TUDelft models
\cite{fiorina_modelling_2014}. The design is a 2-D axisymmetric representation
of the \gls{MSFR} with the sixteen individual outer loops homogenized into
a single outer loop as shown in Figure \ref{fig:msfrgeom}. For the multigroup
group constants calculations in Serpent, the 2-D axisymmetric model is
extended into a 3-D model by a 360-degree rotation about the
central axis. The material definitions are the same as those specified in the
reference \gls{MSFR} model \cite{brovchenko_neutronic_2019}. Accordingly, the
pump and heat exchanger regions
are assumed to be composed of 100\% fuel salt. While this is an approximation,
the outer loop is shielded from the core by a layer of boron carbide and sees
sufficiently low neutron fluxes that we may neglect its exact neutronic
properties.
%
\begin{figure}[t!] 
	\centering
	\includegraphics[width=0.65\textwidth]{msfr-geom-legend}
	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
	simulations in Serpent. All dimensions are in meters.
	\cite{brovchenko_neutronic_2019}}
	\label{fig:msfrgeom}
\end{figure}

Although the present work uses the same 2-D axisymmetric geometry for
generating group constant data from Serpent, two minor differences exist
between the \gls{MSFR} geometry modeled herein, and the geometry used in the
Polimi and TUDelft models \cite{fiorina_modelling_2014}. First, the reactor
geometry for
Moltres excludes the 2-cm-thick structural material around the blanket tank
that separates the fuel and blanket salts. The thickness is much smaller than
the rest of the regions in the geometry and it complicated mesh generation in
the adjacent regions near the boundaries. The
neutronics results in Chapter \ref{chap:nts} show that the overall $k_{eff}$
and other parameters from Moltres show good agreement with that from Serpent.
It also has no direct impact on the temperature distribution results because
the present work solves for the temperature distribution in the primary loop
with homogeneous Neumann boundary conditions on the fuel salt-to-wall
interface. This approach is for consistency with the Polimi and TUDelft
models.

The second difference pertains to the out-of-core section of the primary loop.
The focus for Moltres development has been on core multiphysics over detailed
out-of-core multiphysics. The Moltres model simulates the outer loop as a 1-D
pipe with a pointwise heat sink to represent the heat exchanger. To simulate
the pumps, a Dirichlet boundary condition for inlet velocity
drives the flow in the central core region of the primary loop. At every
timestep, Moltres converges the core and outer loop calculations using Picard
iterations. This approach shares some similarities with the
geometric multiscale modeling approach by Zanetti et al.
\cite{zanetti_geometric_2015}. Future models could create a better
representation of the primary loop by implementing a whole continuous loop
with pressure increases and drops corresponding to the pumps and heat
exchangers.
